---
title: "Scotland-Sparsely-Populated"
author: "Elliot Meador"
date: "07/01/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview
This file contains analysis for identifying low-level changes in population for use in identifying potential case study areas for RESAS 3.4.4. It combines the Scottish Resilience Indicators dataset with the Scottish Sparsely Populated Areas dataset. 


## Data
1. Sparsely populated data files can be accessed [here](https://www.hutton.ac.uk/research/projects/demographic-change-remote-areas).
2. Jon's Shiny App can be accessed [here](https://mappingrd342.shinyapps.io/mappingtool/).
3. Population estimates can be found at [National Records of Scotland](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates).

```{r Sparsely-Pop, warning=F, message=F, fig.height=10, fig.width=6}
library(tidyverse)
library(janitor)

library(sf)
library(tmap)

library(glue)

post_theme <- function(...){
  theme_minimal()+
  theme(...)
}



sparse_pop <- read_csv('/Users/emeador/OneDrive - SRUC/Data/resilience/RD_3.4.1_sparsely_pop.txt') %>% 
    clean_names()


OutputArea2011 <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/output_areas/OutputArea2011_EoR.shp', 
                          quiet = T)

scotland_border <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/scotland_border.gpkg', 
                           quiet = T)


sparse_pop_sf <- OutputArea2011 %>% 
    rename(oa11code = code) %>% 
    left_join(sparse_pop) %>% 
    mutate(post_pre = str_sub(masterpc, 1, 2), 
           SPA = str_detect(spaflag, '^SPA'), 
          subregion_6 = str_remove_all(subregion_6, 'SPA'), 
          subregion_6 = str_squish(subregion_6))


sparse_only <- sparse_pop_sf %>% 
  filter(SPA == T) %>% 
  group_by(subregion_6) %>% 
  summarise()

sparsely_pop_map <- 
  tm_shape(scotland_border)+
   tm_borders()+
  tm_shape(sparse_only)+
    tm_fill('subregion_6', 
            alpha = .85, 
            legend.show = F)+
  tm_text('subregion_6', 
          xmod = 2, 
          ymod=2)+
  tm_layout(title = 'Scottish SPA and its sub-regions (reproduced)')
    
tmap_save(sparsely_pop_map, 
          filename = 'pdf/maps/sparsely_pop_map.pdf', 
          width = 7, 
          height = 11)

sparsely_pop_map
```


```{r}

scotland_pop <- read_csv('/Users/emeador/OneDrive - SRUC/Data/population/small_area/scotland_small_area_2005_2018_tidy.csv')

skye_pop_change <- scotland_pop %>% 
  select(data_zone, data_zone_name, year, population) %>% 
  filter(str_detect(data_zone_name, 'Skye')) %>% 
   arrange(data_zone, year) %>% 
  group_by(data_zone) %>% 
  mutate(base_year = case_when(
    year == min(year) ~ population, 
    T ~ NA_real_
  )) %>% 
  fill(base_year, .direction = 'down') %>% 
  mutate(diff = population - base_year)




skye_pop_change %>% 
  ggplot(aes(year, diff, color = data_zone))+
  geom_hline(yintercept = 0)+
  geom_line(show.legend = F)+
  geom_point(show.legend = F)+
  scale_x_continuous(breaks = 2005:2018)+
  post_theme(panel.grid = element_blank(), 
             text = element_text(family = 'serif'), 
             plot.margin = margin(.5, .5, .5, .5, 'cm'))+
  labs(title = 'Ilse of Skye - Low level population change overtime 2005 - 2018', 
       subtitle = ' SOURCE: National Records of Scotland*', 
       y = 'Change from 2005 (# of people)', 
       caption = 'Each line represents a 2011 datazone, and all populations are aggregated to 2011 datazones.\n*Data are from mid-year estimates.')
```











