---
title: "Scotland-Sparsely-Populated"
author: "Elliot Meador"
date: "07/01/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview
This file contains analysis for identifying low-level changes in population for use in identifying potential case study areas for RESAS 3.4.4. It combines the Scottish Resilience Indicators dataset with the Scottish Sparsely Populated Areas dataset. 


## Data
1. Sparsely populated data files can be accessed [here](https://www.hutton.ac.uk/research/projects/demographic-change-remote-areas).
2. Jon's Shiny App can be accessed [here](https://mappingrd342.shinyapps.io/mappingtool/).
3. Population estimates can be found at [National Records of Scotland](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates).

```{r Sparsely-Pop, warning=F, message=F, fig.height=10, fig.width=6}
library(tidyverse)
library(janitor)

library(sf)
library(tmap)

sparse_pop <- read_csv('/Users/emeador/OneDrive - SRUC/Data/resilience/RD_3.4.1_sparsely_pop.txt') %>% 
    clean_names()


OutputArea2011 <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/output_areas/OutputArea2011_EoR.shp', 
                          quiet = T)

scotland_border <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/scotland_border.gpkg', 
                           quiet = T)


sparse_pop_sf <- OutputArea2011 %>% 
    rename(oa11code = code) %>% 
    left_join(sparse_pop) %>% 
    mutate(post_pre = str_sub(masterpc, 1, 2), 
           SPA = str_detect(spaflag, '^SPA'), 
          subregion_6 = str_remove_all(subregion_6, 'SPA'), 
          subregion_6 = str_squish(subregion_6))


sparse_only <- sparse_pop_sf %>% 
  filter(SPA == T) %>% 
  group_by(subregion_6) %>% 
  summarise()



tm_shape(scotland_border)+
  tm_borders()+
tm_shape(sparse_only)+
    tm_fill('subregion_6', 
            alpha = .85, 
            legend.show = F)+
  tm_text('subregion_6', 
          xmod = 2, 
          ymod=2)+
  tm_layout(title = 'Scottish SPA and its sub-regions (reproduced)')
    
```

