---
title: "Scotland-Sparsely-Populated"
author: "Elliot Meador"
date: "07/01/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Overview
This file contains analysis for identifying low-level changes in population for use in identifying potential case study areas for RESAS 3.4.4. It combines the Scottish Resilience Indicators dataset with the Scottish Sparsely Populated Areas dataset. 


## Data
1. Sparsely populated data files can be accessed [here](https://www.hutton.ac.uk/research/projects/demographic-change-remote-areas).
2. Jon's Shiny App can be accessed [here](https://mappingrd342.shinyapps.io/mappingtool/).
3. Population estimates can be found at [National Records of Scotland](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/population/population-estimates).

```{r Sparsely-Pop, warning=F, message=F, fig.height=10, fig.width=6}
library(tidyverse)
library(janitor)
library(patchwork)

library(sf)
library(tmap)
library(ggrepel)
library(glue)
library(paletteer)

post_theme <- function(...){
  theme_minimal()+
  theme(
    plot.caption = element_text(hjust = 0, face= "italic"), 
    plot.title.position = "plot",  
    plot.caption.position =  "plot",
    ...)
}



sparse_pop <- read_csv('/Users/emeador/OneDrive - SRUC/Data/resilience/RD_3.4.1_sparsely_pop.txt') %>% 
    clean_names()


OutputArea2011 <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/output_areas/OutputArea2011_EoR.shp', 
                          quiet = T)

scotland_border <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/scotland_border.gpkg', 
                           quiet = T)


sparse_pop_sf <- OutputArea2011 %>% 
    rename(oa11code = code) %>% 
    left_join(sparse_pop) %>% 
    mutate(post_pre = str_sub(masterpc, 1, 2), 
           SPA = str_detect(spaflag, '^SPA'), 
          subregion_6 = str_remove_all(subregion_6, 'SPA'), 
          subregion_6 = str_squish(subregion_6))


sparse_only <- sparse_pop_sf %>% 
  filter(SPA == T) %>% 
  group_by(subregion_6) %>% 
  summarise()

sparsely_pop_map <- 
  tm_shape(scotland_border)+
   tm_borders()+
  tm_shape(sparse_only)+
    tm_fill('subregion_6', 
            alpha = .85, 
            legend.show = F)+
  tm_text('subregion_6', 
          xmod = 2, 
          ymod=2)+
  tm_layout(title = 'Scottish SPA and its sub-regions (reproduced)')
    
tmap_save(sparsely_pop_map, 
          filename = 'pdf/maps/sparsely_pop_map.pdf', 
          width = 7, 
          height = 11)

sparsely_pop_map
```


```{r}

scotland_pop <- read_csv('/Users/emeador/OneDrive - SRUC/Data/population/small_area/scotland_small_area_2005_2018_tidy.csv')

skye_pop_change <- scotland_pop %>% 
  select(data_zone, data_zone_name, year, population) %>% 
  filter(str_detect(data_zone_name, 'Skye')) %>% 
   arrange(data_zone, year) %>% 
  group_by(data_zone) %>% 
  mutate(base_year = case_when(
    year == min(year) ~ population, 
    T ~ NA_real_
  )) %>% 
  fill(base_year, .direction = 'down') %>% 
  mutate(diff = population - base_year)

paletteer::palettes_d_names %>% 
  as_tibble() %>% 
  View()

cols <- paletteer_d("RColorBrewer::Accent")

data_zone_cols <- colorRampPalette(
  cols
  )(length(2005:2018))

names(data_zone_cols) <-
  skye_pop_change %>% 
  distinct(data_zone) %>% 
  pull()



(skye_pop_gg <- skye_pop_change %>% 
  ungroup() %>% 
  mutate(data_zone = 
           fct_reorder(data_zone, 
                       diff, 
                       tail, 
                       n = 1, 
                       .desc = TRUE)) %>% 
  ggplot(aes(year, diff, color = data_zone))+
  geom_hline(yintercept = 0)+
  geom_line(size = 1.5, 
            show.legend = T)+
  geom_point(size = 4.5, 
             show.legend = T)+
  scale_x_continuous(breaks = 2005:2018)+
  scale_color_manual(values = data_zone_cols, 
                     name = 'Datazone name')+
  post_theme(panel.grid = element_blank(), 
             text = element_text(family = 'serif'), 
             plot.margin = margin(.5, .5, .5, .5, 'cm'))+
  labs(title = 'Ilse of Skye - Low level population change overtime 2005 - 2018', 
       subtitle = ' SOURCE: National Records of Scotland*', 
       x = 'Year', 
       y = 'Change from 2005 (# of people)', 
       caption = 'Each line represents a 2011 datazone, and all populations are aggregated to 2011 datazones.\n*Data are from mid-year estimates.' ))

ggsave(skye_pop_gg, filename = '/Users/emeador/Documents/R/Scotland_resilience_analysis/png/skye_pop_change.png', width = 8, height = 8, dpi = 500)

system('open /Users/emeador/Documents/R/Scotland_resilience_analysis/png/skye_pop_change.png')


scotland_data_zone_2011 <- st_read('/Users/emeador/OneDrive - SRUC/Data/geographic/boundaries/scotland_data_zone_2011.gpkg')


sky_geo <- scotland_data_zone_2011 %>% 
  filter(str_detect(data_zone_name, 'Skye')) %>% 
  mutate(data_zone = as.factor(data_zone))


skye_map <- ggplot(sky_geo)+
  geom_sf(aes(fill = data_zone), 
          show.legend = F)+
    geom_sf_label(aes(label = data_zone), 
                  size = 1.75, 
                  position = position_dodge(width = .1)) +
  scale_fill_manual(values = data_zone_cols)+
  post_theme(panel.grid = element_blank(), 
             text = element_text(family = 'serif'), 
             plot.margin = margin(.5, .5, .5, .5, 'cm'))+
  labs(title = 'Skye 2011 Datazone Map', 
       subtitle = 'SOURCE: Scottish Government', 
       caption = 'Colours correspond to Ilse of Skye - Low level population change overtime 2005 - 2018 figure')

ggsave(skye_map, filename = '/Users/emeador/Documents/R/Scotland_resilience_analysis/png/skye_map.png', 
       width = 8, 
       height = 8, 
       dpi = 500)

system('open /Users/emeador/Documents/R/Scotland_resilience_analysis/png/skye_map.png')

skye_map + 
  skye_pop_gg +
  plot_annotation(title = 'Skye population change and area')

skye_pop_change_2018 <- skye_pop_change %>% 
  filter(year == max(year)) %>% 
  arrange(data_zone, year) %>% 
  select(data_zone, diff) 
  

sky_geo_change <- sky_geo %>% 
  left_join(skye_pop_change_2018, 
            by = 'data_zone')
  
skye_diff_map <- ggplot(sky_geo_change)+
  geom_sf(aes(fill = diff, 
              color = diff), 
          show.legend = T)+
  scale_fill_gradientn(colors = paletteer_d("RColorBrewer::Spectral"), 
                       name = 'Change from\n 2005'
  )+
  scale_color_gradientn(colors = paletteer_d("RColorBrewer::Spectral")
  , guide = guide_colorbar())+
  post_theme(panel.grid = element_blank())+
  labs(title = 'Ilse of Skye - Population change', 
       subtitle = 'SOURCE: National Records of Scotland', 
       caption = 'Most population change happens in and around Portree.')

ggsave(skye_diff_map, 
       filename = '/Users/emeador/Documents/R/Scotland_resilience_analysis/png/maps/skye_diff_map.png', 
       width = 6, height = 6, dpi = 500)

system('open /Users/emeador/Documents/R/Scotland_resilience_analysis/png/maps/skye_diff_map.png')

```










